Devito for large scale elastic modelling and anisotropic inversion

### 2020-04-21 paper.tex
Add tti-so8-unoptimized.c to the text

One of the main applications of seismic finite-difference modeling in
exploration geophysics is reverse-time migration (RTM), a wave-equation based
seismic imaging technique. Real-world seismic imaging presents a number of
challenges that make applying this method to industry-scale problem sizes
difficult. Firstly, RTM requires an accurate representation of the underlying
physics via sophisticated wave-equations such as the tilted-transverse isotropic
(TTI) wave-equation, for which both forward and adjoint implementations must to
be provided. Secondly, wave-equations must be solved for a large number of
independent experiments, where each individual PDE solve is itself expensive in
terms of FLOPs and memory usage. For certain workloads, limited domain
decomposition, that balances the domain size and the number of independent
experiments, and checkpointing techniques must to be adopted. In the following
sections, we describe an industry-scale seismic imaging problem that poses all
the aforementioned challenges, its implementation with Devito, and the results
of an experiment carried out on the Azure Cloud using a synthetic data set.

@@

Note that while the adjoint of the finite-difference stencil is enabled via the
standard Python \texttt{.T} shortcut, the expression needs to be reordered by hand
since the tilt and azymuth angles are spatially dependent and require to be
inside the second pass of first-order derivative. We can see from these simple
two lines that the rotated stencil involves all second-order derivatives
(\texttt{.dx.dx}, \texttt{.dy.dy} and \texttt{.dz.dz}) and all second-order
cross-derivatives (\texttt{dx.dy}, \texttt{.dx.dz} and \texttt{.dy.dz}) which
leads to a denser stencil support and higher computational complexity
(c.f. \cite{louboutin2016ppf}). For illustrative purposes, the complete
generated code for tti modeling with and without MPI is made available
at \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.

Owing to the very high number of floating-point operations (FLOPs) needed per
grid point for the weighted rotated Laplacian, this anisotropic wave-equation is
extremely challenging to implement. As we show in Table Table~\ref{ttiFLOPs},
and previously analysed in \cite{louboutin2016ppf}, the computational cost with
high-order finite-difference is in the order of thousands of FLOPs per grid
point without optimizations. The version without FLOP-reducing optimizations is
a direct translation of the discretized operators into stencil expressions. The
version with optimizations employs transformations such as common
sub-expressions elimination, factorization, and cross-iteration redundancy
elimination -- the latter being key in removing redundancies introduced by mixed
derivatives. Implementing all of these techniques manually is inherently
difficult and laborious. Further, to obtain the desired performance improvements
it is necessary to orchestrate them with aggressive loop fusion (for data
locality), tiling (for data locality and tensor temporaries), and potentially
ad-hoc vectorization strategies (if rotating registers are used). While an
explanation of the optimization strategies employed by
\href{https://github.com/devitocodes/devito}{Devito} is beyond the scope of this
paper (see \cite{devito-compiler} for details), what is emphasized here is that
users can easily take full advantage of these optimization without needed to
concern themselves with the details.

>>>>

One of the main applications of seismic finite-difference modeling in
exploration geophysics is reverse-time migration (RTM), a wave-equation based
seismic imaging technique. Real-world seismic imaging presents a number of
challenges that make applying this method to industry-scale problem sizes
difficult. Firstly, RTM requires an accurate representation of the underlying
physics via sophisticated wave-equations such as the tilted-transverse isotropic
(TTI) wave-equation, for which both forward and adjoint implementations must to
be provided. Secondly, wave-equations must be solved for a large number of
independent experiments, where each individual PDE solve is itself expensive in
terms of FLOPs and memory usage. For certain workloads, limited domain
decomposition, which balances the domain size and the number of independent
experiments, as well as checkpointing techniques must be adopted. In the following
sections, we describe an industry-scale seismic imaging problem that poses all
the aforementioned challenges, its implementation with Devito, and the results
of an experiment carried out on the Azure Cloud using a synthetic data set.

@@

Note that while the adjoint of the finite-difference stencil is enabled via the
standard Python \texttt{.T} shortcut, the expression needs to be reordered by hand
since the tilt and azymuth angles are spatially dependent and require to be
inside the second pass of first-order derivative. We can see from these simple
two lines that the rotated stencil involves all second-order derivatives
(\texttt{.dx.dx}, \texttt{.dy.dy} and \texttt{.dz.dz}) and all second-order
cross-derivatives (\texttt{dx.dy}, \texttt{.dx.dz} and \texttt{.dy.dz}) which
leads to a denser stencil support and higher computational complexity
(c.f. \cite{louboutin2016ppf}). For illustrative purposes, the complete
generated code for tti modeling with and without MPI is made available
at \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{tti-so8-unoptimized.c}, \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.

Owing to the very high number of floating-point operations (FLOPs) needed per
grid point for the weighted rotated Laplacian, this anisotropic wave-equation
is extremely challenging to implement. As we show in Table~\ref{ttiFLOPs}, and
previously analysed in \cite{louboutin2016ppf}, the computational cost with
high-order finite-difference is in the order of thousands of FLOPs per grid
point without optimizations. The version without FLOP-reducing optimizations is
a direct translation of the discretized operators into stencil expressions (see
\texttt{tti-so8-unoptimized.c}). The version with optimizations employs
transformations such as common sub-expressions elimination, factorization, and
cross-iteration redundancy elimination -- the latter being key in removing
redundancies introduced by mixed derivatives. Implementing all of these
techniques manually is inherently difficult and laborious. Further, to obtain
the desired performance improvements it is necessary to orchestrate them with
aggressive loop fusion (for data locality), tiling (for data locality and
tensor temporaries), and potentially ad-hoc vectorization strategies (if
rotating registers are used). While an explanation of the optimization
strategies employed by \href{https://github.com/devitocodes/devito}{Devito} is
beyond the scope of this paper (see \cite{devito-compiler} for details), what
is emphasized here is that users can easily take full advantage of these
optimization without needed to concern themselves with the details.
